\toolnameend: A Tool for Visualizing Patterns of Spreadsheet Function Combinations

### 2016-08-13 doc/paper.tex
Walkthrough work

\usepackage{xspace}
\usepackage{subfigure}

@@

	Other tools focus on creating visualizations external to the sheets. Hermans
	and colleagues address a spreadsheet programmer's information needs by making
	dataflow diagrams from individual sheets through the tools
	GyroSAT~\cite{hermans2011supporting} and Breviz~\cite{hermans2011breviz}.
	\toolname is among this type, creating standalone visualizations about
	spreadsheets instead of drawing connections inside the spreadsheet itself. In
	contrast to the other visualization work, our work focuses on representing
	functions and how they combine into complex formulae, rather than exploring the
	data structures of any particular spreadsheet.  \par

	Many other tools inspect the structure of formulae but from the
	perspective of finding bad design.  Following Fowler's initial description of
	code smells in object-oriented design~\cite{fowler2009refactoring}, many
	researchers have applied similar observations to spreadsheets and have
	cataloged classes of problematic
	formulae~\cite{hermans2012detecting}~\cite{cunha2012towards}~\cite{asavametha2012detecting}, often accompanying them with detection tools~\cite{abreu2014smelling} and investigations into real datasets~\cite{jansen2015code}. For example, Hermans, leveraging her work with the aforementioned dataflow diagrams, has worked with assessing common inter-worksheet smells, such as feature envy and inappropriate intimacy, wherein references between worksheets (rather than cells) in a file suggest a problem~\cite{hermans2012detectinginter}. Other tools, like Badame's Excel refactoring~\cite{badame2012refactoring}, are concerned more with understanding the formula in service of changing them. Though \toolname supports the search for bad design, as we will see, the tool itself takes a much more agnostic approach in presentation, offering examples without regard for design quality. \par

@@

	\subsection{Walkthrough of the Tool} \label{sec:walkthrough}Before discussing
	the tool's development, we will give a basic example of how to use
	the tool. For this, we will work from the visualization built on the Enron
	corpus to answer this question: \textit{ What kinds of functions do people use
		to define the condition in an IF function?} \par

	In Excel, the IF functions is one of the logical
	functions: given a condition as the first argument that evaluates to either
	true or false, it takes on the value of either the second or third argument,
	respectively. Or, as the documentation writes out, ``IF(Something is True, then
	do something, otherwise do something
	else)"\footnote{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2?ui=en-US\&rs=en-US\&ad=US}. Furthermore, the third argument is optional; if left out, the function will return a default ``FALSE" value if necessary. Examples of this function in practice can be seen in Figure~\ref{fig:ifexample}.

@@

	When the user first approaches the tree for a given top-level function -- that
	is, a function nested within none other -- only a few nodes are visible, as
	shown in Figure~\ref{fig:startpic}. \par

@@

	The visualization so far comprises two types of nodes: the circle
	\includegraphics{glossary-greenonly}, which represents a discrete function in
	the formula and is sized according to its frequency relative to the number of
	times the root function appeared; and the numbered squares
	\includegraphics{glossary-blue}, which represent the positions of arguments
	within its parent function. From this, we can confirm that, of the times it was
	observed, IF can have at most three arguments passed into it, which corresponds
	with its specifications in the API. \par

	Knowing that it is the first argument which contains the conditionals, we click
	on the square labeled ``1" to explore. To save space, when the tool finds more
	than 10 unique arguments in a position, it displays only the first ten, with an
	option to display the rest by clicking the arrow
	\includegraphics{glossary-arrow}.

@@

	We see that IF contains a number of comparison operators as its first argument,
	such as = and $\le$, and boolean-returning functions, like ISNA and AND, with
	simple equality being the most common and ABS being the least. From here, we
	can further explore the common options among these functions. Clicking on the
	``=" node will yield two arguments and expanding each of them will peer into
	the range of common values of equality comparison, which Figure
	\ref{fig:fullpic} shows. \par

@@

	For both sides of the equal sign, the operator has certain types of arguments
	which predominate over the others: on the left side is most often a reference
	to another cell; on the right, a string or number literal, which makes sense
	for the case of confirming a value in another cell before assigning this one.
	Furthermore, a tooltip accompanies each function node in the tree, providing a
	concrete example of a function that uses this structure. If this single
	instance is not enough, then the user also has the option to double-click the
	individual function node, which will open up a new tab with a table of many
	more examples from dataset.

>>>>

\usepackage{xspace}
\usepackage{subfigure}
\usepackage{url}

@@

	Other tools focus on creating visualizations external to the sheets. Hermans
	and colleagues address a spreadsheet programmer's information needs by making
	dataflow diagrams from individual sheets through the tools
	GyroSAT~\cite{hermans2011supporting} and Breviz~\cite{hermans2011breviz}.
	\toolname is among this type, creating standalone visualizations about
	spreadsheets instead of drawing connections inside the spreadsheet itself. In
	contrast to the other visualization work, our work focuses on representing
	functions and how they combine into complex formulae, rather than exploring the
	data structures of any particular spreadsheet.

	Many other tools inspect the structure of formulae but from the
	perspective of finding bad design.  Following Fowler's initial description of
	code smells in object-oriented design~\cite{fowler2009refactoring}, many
	researchers have applied similar observations to spreadsheets and have
	cataloged classes of problematic
	formulae~\cite{hermans2012detecting}~\cite{cunha2012towards}~\cite{asavametha2012detecting}, often accompanying them with detection tools~\cite{abreu2014smelling} and investigations into real datasets~\cite{jansen2015code}. For example, Hermans, leveraging her work with the aforementioned dataflow diagrams, has worked with assessing common inter-worksheet smells, such as feature envy and inappropriate intimacy, wherein references between worksheets (rather than cells) in a file suggest a problem~\cite{hermans2012detectinginter}. Other tools, like Badame's Excel refactoring~\cite{badame2012refactoring}, are concerned more with understanding the formula in service of changing them. Though \toolname supports the search for bad design, as we will see, the tool itself takes a much more agnostic approach in presentation, offering examples without regard for design quality.

@@

	\subsection{Walkthrough of the Tool} \label{sec:walkthrough}Before discussing
	\toolnameposs development, we will give a basic example of how to use
	it. We will work from the visualization built on the Enron
	corpus to answer this question: \textit{ What kinds of functions do people use
		to define the condition in an IF function?}

	In Excel, the IF functions is one of the logical functions: given a condition
	as the first argument that evaluates to either true or false, it takes on the
	value of either the second or third argument respectively. Or, as the
	documentation writes out, ``IF(Something is True, then do something, otherwise
	do something
	else)"\footnote{\url{https://support.office.com/en-us/article/IF-function-69aed7c9-4e8a-4755-a9bc-aa8bbff73be2?ui=en-US\&rs=en-US\&ad=US}}. Furthermore, the third argument is optional; if left out, the function will return a default ``FALSE" value if necessary. Examples of this function in practice can be seen in Figure~\ref{fig:ifexample}.

@@

	When the user first approaches the tree for a given top-level function -- that
	is, a function nested within none other -- only a few nodes are visible, as
	shown in Figure~\ref{fig:startpic}.

@@

	The visualization so far comprises two types of nodes. The first are the
	circles \includegraphics{glossary-greenonly}, which represens a discrete
	function in the formula. These are sized according to their frequency relative
	to the number of times the root function appeared. The second kind of node are
	the numbered squares \includegraphics{glossary-blue}, which represent the
	positions of arguments within its parent function. Looking back at the tree, we
	can confirm that IF functions in this dataset had at most three arguments
	passed into them, which corresponds with its specifications in the API.

	Knowing that the first argument of IF contains the conditionals, we click on
	the square labeled ``1" to explore. When the tool finds more than 10 unique
	arguments in a position, it saves space by displaying only the first ten and
	includes an option to display the rest by clicking the arrow
	\includegraphics{glossary-arrow}.

@@

	In Figure~\ref{fig:expandif}, we see that IF's first arguments are most often comparison operators,
	such as = and $\le$, and boolean-returning functions, like ISNA and AND. Of these,
	simple equality is the most common and ABS is the least. From here, we
	can further explore the common options among these functions. Clicking on the
	``=" node will yield two arguments and expanding each of them will show
	the common values on either side of the equals sign, as in Figure~\ref{fig:fullpic}.

@@

	The equality operator has certain types of arguments
	which predominate over the others: on the left side is most often a reference
	to another cell; on the right is a string or number literal, which makes sense
	for the case of confirming a value in another cell before assigning this one.
	Furthermore, a tooltip accompanies each function node in the tree, providing a
	concrete example of a function that uses this structure. If this single
	instance is not enough, then the user also has the option to open up a new
	window with a table of many more examples from dataset.
